Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016

### 2019-04-21 draft-paper.Rmd
Updated Jeremy's email address

---
title: "Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016"
author:
- familyname: Forbes
  othernames: Jeremy
  address: Monash University\newline Melbourne, Australia
  email: jeremy.forbes@monash.edu
  correspondingauthor: true
- familyname: Cook
  othernames: Dianne
  address: Monash University\newline Melbourne, Australia
- familyname: Hyndman
  othernames: Rob J
  address: Monash University\newline Melbourne, Australia
abstract: "We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016. Socio-demographic information is derived from the Australian Census, which occurs every five years. Since a Census is not directly available for each election, spatio-temporal imputation is employed to estimate Census data for the electorates at the time of each election. This accounts for both spatial and temporal changes in electoral characteristics between Censuses. To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can be thought of as the unobserved political climate in each electorate. Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections. Education and unemployment are amongst those that have varying effects. It is also found that between 2004 and 2013, the spatial effect was insignificant, meaning that electorates voted effectively independently. All data featured in this study has been contributed to the `eechidna` `R` package (available on CRAN)."
keywords: "federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna"
wpnumber:
jelcodes: C31, C33, D72
blind: false
cover: false
toc: false
bibliography: references.bib
biblio-style: authoryear-comp
header-includes: \usepackage{float}
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---

>>>>

---
title: "Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016"
author:
- familyname: Forbes
  othernames: Jeremy
  address: Monash University\newline Melbourne, Australia
  email: jeremyforbes1995@gmail.com
  correspondingauthor: true
- familyname: Cook
  othernames: Dianne
  address: Monash University\newline Melbourne, Australia
- familyname: Hyndman
  othernames: Rob J
  address: Monash University\newline Melbourne, Australia
abstract: "We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016. Socio-demographic information is derived from the Australian Census, which occurs every five years. Since a Census is not directly available for each election, spatio-temporal imputation is employed to estimate Census data for the electorates at the time of each election. This accounts for both spatial and temporal changes in electoral characteristics between Censuses. To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can be thought of as the unobserved political climate in each electorate. Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections. Education and unemployment are amongst those that have varying effects. It is also found that between 2004 and 2013, the spatial effect was insignificant, meaning that electorates voted effectively independently. All data featured in this study has been contributed to the `eechidna` `R` package (available on CRAN)."
keywords: "federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna"
wpnumber:
jelcodes: C31, C33, D72
blind: false
cover: false
toc: false
bibliography: references.bib
biblio-style: authoryear-comp
header-includes: \usepackage{float}
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---
